---
title: "Homework 3"
author: "PSTAT 131/231"
output:
    html_document:
      toc: true
      toc_float: true
      code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
library(tidymodels)
library(ISLR) 
library(ISLR2) 
library(discrim)
library(poissonreg)
library(corrr)
library(klaR) 
Titanic <- read.csv("C:\\Users\\Jonat\\OneDrive\\schoolwork\\PSTAT 131\\HW\\HW 3\\homework-3\\data\\titanic.csv")
```

### Question 1

Split the data, stratifying on the outcome variable, `survived.`  You should choose the proportions to split the data into. Verify that the training and testing data sets have the appropriate number of observations. Take a look at the training data and note any potential issues, such as missing data.

```{r}
set.seed(3435)

sum(is.na(Titanic)) #There are 866 spots of missing data.

Titanic$survived <-factor(Titanic$survived, levels=c("Yes", "No"))
Titanic$pclass <-factor(Titanic$pclass)
Titanic %>% head()

Titanic_split <- initial_split(Titanic, prop = 0.70, strata = survived)
Titanic_train <- training(Titanic_split)
Titanic_test <- testing(Titanic_split)
Titanic_train %>% head()
```

Why is it a good idea to use stratified sampling for this data?
- I think it is a good idea to use stratified sampling because we can divide them into mutually exclusive groups and subgroups that can possibly take on different mean values for the variable Survived. We will also be able to obtain estimates with a relatively low variance. We also have a large sample.

### Question 2

Using the **training** data set, explore/describe the distribution of the outcome variable `survived`.

```{r}
library(ggplot2)
ggplot(Titanic_train,aes(survived)) + geom_bar() + xlab("Survived")

```
Based off the graph I could say that more people did not survive. There was about 270 people who survived according to this data set. The are more than 450 people who did not survive.

### Question 3

Using the **training** data set, create a correlation matrix of all continuous variables. Create a visualization of the matrix, and describe any patterns you see. Are any predictors correlated with each other? Which ones, and in which direction?
```{r}
cor_Titanic <- Titanic_train %>%
  select_if(is.numeric) %>%
  correlate()
rplot(cor_Titanic)
cor_Titanic %>%
  stretch() %>%
  ggplot(aes(x, y, fill = r)) +
  geom_tile() +
  geom_text(aes(label = as.character(fashion(r))))
```
The only variables that seem to have an OK correlation is the relation between the variables sib_sp (# of siblings / spouses aboard the Titanic) and parch (# of parents / children aboard the Titanic). They have a positive correlation of 0.43 or 43%. Everything else does not have a strong enough correlation to recognize.


### Question 4

Using the **training** data, create a recipe predicting the outcome variable `survived`. Include the following predictors: ticket class, sex, age, number of siblings or spouses aboard, number of parents or children aboard, and passenger fare.

Recall that there were missing values for `age`. To deal with this, add an imputation step using `step_impute_linear()`. Next, use `step_dummy()` to **dummy** encode categorical predictors. Finally, include interactions between:

-   Sex and passenger fare, and
-   Age and passenger fare.

You'll need to investigate the `tidymodels` documentation to find the appropriate step functions to use.
```{r}
Titanic_recipe <- recipe(survived ~ pclass + sex + age + sib_sp + parch + fare, data = Titanic_train) %>%
  step_impute_linear(age) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_interact(terms = ~ starts_with("sex"):fare) %>%
  step_interact(terms = ~ age:fare)
Titanic_recipe
```

### Question 5

Specify a **logistic regression** model for classification using the `"glm"` engine. Then create a workflow. Add your model and the appropriate recipe. Finally, use `fit()` to apply your workflow to the **training** data.

***Hint: Make sure to store the results of `fit()`. You'll need them later on.***
```{r}
lm_reg <- logistic_reg() %>% 
  set_engine("glm") %>% 
  set_mode("classification")
lm_workflow <- workflow() %>% 
  add_model(lm_reg) %>% 
  add_recipe(Titanic_recipe)
lm_fit <- fit(lm_workflow, Titanic_train)
```

### Question 6

**Repeat Question 5**, but this time specify a linear discriminant analysis model for classification using the `"MASS"` engine.
```{r}
lda_mod <- discrim_linear() %>% 
  set_mode("classification") %>% 
  set_engine("MASS")
lda_wkflow <- workflow() %>% 
  add_model(lda_mod) %>% 
  add_recipe(Titanic_recipe)
lda_fit <- fit(lda_wkflow, Titanic_train)
```

### Question 7

**Repeat Question 5**, but this time specify a quadratic discriminant analysis model for classification using the `"MASS"` engine.
```{r}
qda_model <- discrim_quad()  %>% 
  set_mode("classification")%>% 
  set_engine("MASS")
qda_workflow <- workflow() %>% 
  add_model(qda_model) %>% 
  add_recipe(Titanic_recipe)
qda_fit <- fit(qda_workflow, Titanic_train)
```

### Question 8

**Repeat Question 5**, but this time specify a naive Bayes model for classification using the `"klaR"` engine. Set the `usekernel` argument to `FALSE`.

```{r}
library(klaR)
nb_model <- naive_Bayes()  %>% 
  set_mode("classification")%>% 
  set_engine("klaR")  %>% 
  set_args(usekernel = FALSE)
nb_workflow <- workflow() %>% 
  add_model(nb_model) %>% 
  add_recipe(Titanic_recipe)
nb_fit <- fit(nb_workflow, Titanic_train)
```
### Question 9

Now you've fit four different models to your training data.

Use `predict()` and `bind_cols()` to generate predictions using each of these 4 models and your **training** data. Then use the *accuracy* metric to assess the performance of each of the four models.

Which model achieved the highest accuracy on the training data?
```{r}
library(dplyr)
# logistic regression
lm_predict <- bind_cols(predict(lm_fit, new_data=Titanic_train), Titanic_train %>% dplyr::select(survived))
lm_predict
lm_accuracy <- lm_predict %>%
          accuracy(truth=survived, estimate = .pred_class)
lm_accuracy
# LDA
lda_predict <- bind_cols(predict(lda_fit, new_data=Titanic_train), Titanic_train %>% dplyr::select(survived))
lda_predict
lda_accuracy <- lda_predict %>%
          accuracy(truth=survived, estimate = .pred_class)
lda_accuracy
# QDA
qda_predict <- bind_cols(predict(qda_fit, new_data = Titanic_train), Titanic_train %>% dplyr::select(survived))
qda_predict
qda_accuracy <- qda_predict %>%
          accuracy(truth=survived, estimate = .pred_class)
qda_accuracy
# Naive Bayes
nb_predict <- suppressWarnings(bind_cols(predict(nb_fit, new_data = Titanic_train), Titanic_train %>% dplyr::select(survived)))
nb_predict
nb_accuracy <- nb_predict %>%
          accuracy(truth=survived, estimate = .pred_class)
nb_accuracy
```
The logisitic regression is the model that had the most precise accuracy with a 81.4% accuracy.

### Question 10

Fit the model with the highest training accuracy to the **testing** data. Report the accuracy of the model on the **testing** data.

Again using the **testing** data, create a confusion matrix and visualize it. Plot an ROC curve and calculate the area under it (AUC).

How did the model perform? Compare its training and testing accuracies. If the values differ, why do you think this is so?
```{r}
# The highest training accuracy testing data
lm_predict
lm_accuracy
confusion_matrix <- augment(lm_fit, new_data = Titanic_test) %>%
  conf_mat(truth = survived, estimate = .pred_class) %>% 
  autoplot(type = "heatmap")
confusion_matrix
ROC <- augment(lm_fit, new_data = Titanic_test) %>%
  roc_curve(survived, .pred_Yes) %>%
  autoplot()
ROC
AUC <- pROC::auc(augment(lm_fit, new_data = Titanic_test)$survived, augment(lm_fit, new_data = Titanic_test)$.pred_Yes)
AUC
```
The Training and Test model performed similar enough. The value for the training is 0.8138042 and the value for the testing is 0.88. Since they are pretty close we can say that the model has a pretty accurate prediction and should not have the issue of overfitting. Based on our confusion matrix we can also say that it's predictions are mainly accurate.